comgen-outline

### 2017-06-15 paper/comgen-paper.Rmd
incorporate raw generics data file

```{r generic-endorsement-priors}
d.gen.endorse.priors <- read.csv(paste(project.path, 
                                       "data/generics/endorsement/",
                                       "naturalGenerics-prior-trials-n57.csv", sep = ""))
gen.endorse.properties <- levels(d.gen.endorse.priors$Property)

@@

@@

d.gen.endorse <- read.csv(paste(project.path, "data/generics/endorsement/",
                    "truth-judgments.csv", sep = ""))

d.gen.endorse.summary <- left_join(
  d.gen.endorse, 
  d.gen.endorse.catch %>% select(workerid, pass)
  ) %>%
  filter(pass == 1) %>%
  rowwise() %>%
  mutate(response = ifelse(response == "agree-key", 1, 0),
         sentence = gsub('&quotechar', '', sentence),
         sentence = gsub('lyme', 'Lyme', sentence)) %>%
  group_by(sentence) %>%
  multi_boot_standard(column = "response") %>%
  ungroup() %>%
  mutate(sentence = factor(sentence, levels = sentence[order(mean)]))

@@

Habituals about people are of particular importance because they are likely central to intuitive theories of personality and behavior. 
Young children can use actors' behaviors to make inferences about what the actors are like more generally [e.g., @Repacholi1997; @Seiver2013].
@McGuire1986 explored 5th to 12th graders' responses to open ended probes such as ``Tell us about your school / family''. 
A surprisingly large number of utterances (over 85\% in their study) were about people, and roughly two-thirds of those utterances used action verbs (e.g. ``My brother works part-time at the restaurant.'').
Habitual language may be a more conservative form of trait language (e.g. \emph{Bill is a smoker.}) and convey that behaviors are relatively enduring [@Gelman1999; @Gelman2004].

@@

```{r figure-causals-priors, fig.caption = "Four background distributions participants were exposed to in Expt. 4 (stimuli) and priors reconstructed from participants' responses (model). ", fig.width = 8, fig.height = 3}
sample.dists <- bind_rows(
  data.frame(
    dist = "rare weak",
    x = c(rbeta(n = 3000, shape1 = 2, shape2 = 10),
          rbeta(n = 7000, shape1 = 2, shape2 = 50)
          )
  ),
  data.frame(
    dist = "common weak",
    x = rbeta(n = 10000, shape1 = 2, shape2 = 10)
  ),
  data.frame(
    dist = "rare deterministic",
    x = c(rbeta(n = 3000, shape1 = 50, shape2 = 2),
          rbeta(n = 7000, shape1 = 2, shape2 = 50)
          )
  ),
  data.frame(
    dist = "common deterministic",
    x = rbeta(n = 10000, shape1 = 50, shape2 = 2)
  )
) %>%
    mutate(Distribution = factor(dist, 
                               levels = c("common deterministic", "rare deterministic", "common weak", "rare weak"),
         labels = c("Common Deterministic", "Rare Deterministic",
                    "Common Weak", "Rare Weak")))

@@

```{r causal-model-insets, fig.width = 8, fig.height = 3.5}

m.cas.fullmodel.prior.parameters <- m.cas.samp %>%
  filter(type == "prior") %>%
  rename(variable = item, parameter = roundedFreq) %>%
  group_by(dist, variable, parameter) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

@@

causals.endorsement.insets <- cas.listener.predictions %>% 
    mutate(Parameter = factor(Parameter, levels = c("state_Prior",
                                                  "state_Posterior"
                                                  ),
                            labels = c("Listener Prior (Posterior given Silence)",
                                       "Listener Posterior given Causal")),
    Distribution = factor(distribution, 
                               levels = c("common_deterministic", "rare_deterministic", "common_weak", "rare_weak"),
         labels = c("Common Deterministic", "Rare Deterministic",
                    "Common Weak", "Rare Weak"))) %>%
  ggplot(., aes( x = value, fill = Distribution, color = Distribution, lty = Parameter, alpha = Parameter ))+
  geom_density(aes(y = ..scaled..), adjust = 4, size = 1)+
  facet_wrap(~Distribution, nrow = 1)+
  # geom_text_repel(data = category.text.labels,
  #                 aes(label = category, x = x , y = y),
  #                 inherit.aes = F, color = "#2b83ba")+
  scale_fill_solarized()+ scale_color_solarized()+
  #scale_fill_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
  #scale_color_manual(values = c("#636363", "#abdda4", "#2b83ba", "#d7191c"))+
  #scale_fill_manual(values = c("#636363", "#d7191c"))+#, "#2b83ba"))+
  #scale_color_manual(values = c("#636363", "#d7191c"))+#, "#2b83ba"))+
  scale_alpha_manual(values = c(0.1, 0.8))+#, 0))+
  #scale_linetype_manual(values = c(3, 4, 2, 1))+
  scale_linetype_manual(values = c(2, 1))+
  scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
  scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
  xlab("Prevalence") +
  ylab("Scaled probability density")+
  theme(legend.position = "bottom", legend.title = element_blank())+
  guides(fill = F, color = F)#,
  #       axis.text.x = element_text(angle = 90))+
      # scale_x_continuous(limits = c(-5, 8),
      #                breaks = c(-5, 0, 2.5, 4, 5.9),
      #                labels = c("almost never", "annually", "monthly", "weekly", "daily"))
```

```{r figure-causals-endorsement, fig.width=7, fig.height=3.5}
figure.causals.endorsement <- bind_rows(
  m.cas.fullmodel.endorsement %>%
    mutate(src = "model"),
  d.caus.endorse.summary %>%
    select(distribution, frequency, MAP_h, low, high) %>%
    rename(dist = distribution, MAP = MAP_h, cred_lower = low, cred_upper = high) %>%
    mutate(src = 'data')
) %>%
  ungroup() %>%
  mutate(
    src = factor(src, levels = c( "data", "model")),
    dist = factor(dist, 
           levels = c("common_deterministic", "rare_deterministic", "common_weak", "rare_weak"),
           labels = c("Common Deterministic", "Rare Deterministic",
                      "Common Weak", "Rare Weak"))
    ) %>%
  ggplot(., aes( x = dist, y = MAP, ymin = cred_lower, ymax = cred_upper, fill = dist,
                        group = src, alpha = src))+
  geom_bar(stat='identity', position = position_dodge(), width = 0.3, color = 'black')+
  geom_linerange(position = position_dodge(0.3), alpha = 1)+
  #geom_line(position = position_dodge(0.3))+
  #theme(axis.text.x = element_text(angle = 90))+
  geom_hline(yintercept = 0.5, lty = 2, color = 'black', alpha = 0.4)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 1))+
  scale_alpha_manual(values = c(1, 0.3))+
  scale_fill_solarized()+
  xlab("")+
  ylab("Proportion Causal Endorsement")+
  facet_grid(.~frequency)+
  coord_flip()+
  guides(fill = F)+
  theme(legend.position = 'bottom', legend.direction = 'horizontal', legend.title = element_blank())
```

```{r figure-causals, fig.width=11, fig.height=8}
grid.arrange(figure.causals.endorsement, causals.endorsement.insets, ncol = 1,
             layout_matrix = cbind(c(1,1,2)))

>>>>

```{r generic-endorsement-priors}
# this data set already excludes the 2nd attempt of 3 participants who completed the experiment a second time

d.gen.endorse.priors <- read.csv(paste(project.path, 
                                       "data/generics/endorsement/",
                                       "naturalGenerics-prior-trials-n57.csv", sep = ""))
gen.endorse.properties <- levels(d.gen.endorse.priors$Property)

@@

#### Method
```{r}
d1 <- read.csv(paste(project.path, "data/generics/endorsement/",
                    "truth-judgments.csv", sep = ""))
d2 <- read.csv(paste(project.path, "data/generics/endorsement/",
                    "naturalGenerics-trials-formatted.csv", sep = ""))

head(d1)
head(d2)

write.csv(
  left_join(
  d1 %>%
    mutate(sentence = as.character(sentence),
           sentence = gsub('&quotechar', '', sentence),
         sentence = gsub('lyme', 'Lyme', sentence)),
  unique(d2 %>%
    mutate(sentence = paste(Category, " ", Property, ".", sep = "")) %>%
    select(Category, Property, sentence))
), file = paste(project.path, "data/generics/endorsement/",
                    "truth-judgments-n100.csv", sep = ""), row.names = F)
```

@@

d.gen.endorse <- read.csv(paste(project.path, "data/generics/endorsement/",
                    "truth-judgments-n100.csv", sep = ""))

d.gen.endorse.summary <- left_join(
  d.gen.endorse, 
  d.gen.endorse.catch %>% select(workerid, pass)
  ) %>%
  filter(pass == 1) %>%
  rowwise() %>%
  mutate(response = ifelse(response == "agree-key", 1, 0)) %>%
  group_by(sentence) %>%
  multi_boot_standard(column = "response") %>%
  ungroup() %>%
  mutate(sentence = factor(sentence, levels = sentence[order(mean)]))

@@

Habituals about people are of particular importance because they are likely central to intuitive theories of personality and behavior. 
Young children can use actors' behaviors to make inferences about what the actors are like more generally [e.g., @Repacholi1997; @Seiver2013].
@McGuire1986 explored 5th to 12th graders' responses to open ended probes such as "Tell us about your school / family". 
A surprisingly large number of utterances (over 85\% in their study) were about people, and roughly two-thirds of those utterances used action verbs (e.g., "My brother works part-time at the restaurant.").
Habitual language may be a more conservative form of trait language (e.g., \emph{Bill is a smoker.}) and convey that behaviors are relatively enduring [@Gelman1999; @Gelman2004].

@@

```{r figure-causals-priors, fig.caption = "Four background distributions participants were exposed to in Expt. 4 (stimuli) and priors reconstructed from participants' responses (model). ", fig.width = 8, fig.height = 3, eval = F}
sample.dists <- bind_rows(
  data.frame(
    dist = "rare weak",
    x = c(rbeta(n = 3000, shape1 = 2, shape2 = 10),
          rbeta(n = 7000, shape1 = 2, shape2 = 50)
          )
  ),
  data.frame(
    dist = "common weak",
    x = rbeta(n = 10000, shape1 = 2, shape2 = 10)
  ),
  data.frame(
    dist = "rare deterministic",
    x = c(rbeta(n = 3000, shape1 = 50, shape2 = 2),
          rbeta(n = 7000, shape1 = 2, shape2 = 50)
          )
  ),
  data.frame(
    dist = "common deterministic",
    x = rbeta(n = 10000, shape1 = 50, shape2 = 2)
  )
) %>%
    mutate(Distribution = factor(dist, 
                               levels = c("common deterministic", "rare deterministic", "common weak", "rare weak"),
         labels = c("Common Deterministic", "Rare Deterministic",
                    "Common Weak", "Rare Weak")))

@@

```{r causal-model-insets, fig.width = 8, fig.height = 3.5, cache = F}
m.cas.fullmodel.prior.parameters <- m.cas.samp %>%
  filter(type == "prior") %>%
  rename(variable = item, parameter = roundedFreq) %>%
  group_by(dist, variable, parameter) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

@@

inset.color.order <- c( "#268bd2", "#dc322f","#d33682","#859900")

causals.endorsement.insets <- cas.listener.predictions %>% 
    mutate(Parameter = factor(Parameter, 
                              levels = c("state_Prior","state_Posterior"),
                            labels = c("Listener Prior (Posterior given Silence)",
                                       "Listener Posterior given Causal")),
    Distribution = factor(distribution, 
                               levels = c("common_deterministic", "rare_deterministic", "common_weak", "rare_weak"),
         labels = c("Common Deterministic", "Rare Deterministic",
                    "Common Weak", "Rare Weak"))) %>%
  ggplot(., aes( x = value, fill = Distribution, color = Distribution, lty = Parameter, alpha = Parameter ))+
  geom_density(aes(y = ..scaled..), adjust = 4, size = 1)+
  facet_wrap(~Distribution, nrow = 1)+
  scale_fill_manual(values =inset.color.order )+
  scale_color_manual(values = inset.color.order)+
  scale_alpha_manual(values = c(0.1, 0.8))+
  scale_linetype_manual(values = c(2, 1))+
  scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
  scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
  xlab("Prevalence") +
  ylab("Scaled probability density")+
  theme(legend.position = "bottom", legend.title = element_blank())+
  guides(fill = F, color = F)
```

```{r figure-causals-endorsement, fig.width=7, fig.height=3.5, cache = F}
figure.causals.endorsement <- bind_rows(
  m.cas.fullmodel.endorsement %>%
    mutate(src = "model"),
  d.caus.endorse.summary %>%
    select(distribution, frequency, MAP_h, low, high) %>%
    rename(dist = distribution, MAP = MAP_h, cred_lower = low, cred_upper = high) %>%
    mutate(src = 'data')
) %>%
  ungroup() %>%
  mutate(
    src = factor(src, levels = c( "data", "model")),
    dist = factor(dist, 
           levels = c("rare_weak", "common_weak", "rare_deterministic","common_deterministic"),
           labels = c(
                       "Rare Weak", "Common Weak","Rare Deterministic","Common Deterministic"))
    ) %>%
  ggplot(., aes( x = dist, y = MAP, ymin = cred_lower, ymax = cred_upper, fill = dist,
                        group = src, alpha = src))+
  geom_bar(stat='identity', position = position_dodge(), width = 0.3, color = 'black')+
  geom_linerange(position = position_dodge(0.3), alpha = 1)+
  #geom_line(position = position_dodge(0.3))+
  #theme(axis.text.x = element_text(angle = 90))+
  geom_hline(yintercept = 0.5, lty = 2, color = 'black', alpha = 0.4)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 1))+
  scale_alpha_manual(values = c(1, 0.3))+
  #scale_fill_solarized()+
  scale_fill_manual(values = c("#859900","#d33682","#dc322f", "#268bd2"))+
  xlab("")+
  ylab("Proportion Causal Endorsement")+
  facet_grid(.~frequency)+
  coord_flip()+
  guides(fill = F)+
  theme(legend.position = 'bottom', legend.direction = 'horizontal', legend.title = element_blank())
```

```{r figure-causals, fig.width=11, fig.height=8, cache = F}
grid.arrange(figure.causals.endorsement, causals.endorsement.insets, ncol = 1,
             layout_matrix = cbind(c(1,1,2)))
